---
title: "List 1"
output:
  pdf_document: default
  html_notebook: default
---

## Init and Dataset import

```{r, message=FALSE}
library(tidyverse)
library(jmv)
hand1994_babies <- read_csv("data/hand1994_babies.csv", 
    col_types = cols(Survival = col_factor(levels = c("no", "yes"))))

hand1994_elderheight <- read_lines("data/hand1994_elderheight.csv") %>%
  str_split_fixed(" ", n = Inf) %>%
  as.numeric %>%
  na.omit()

hand1994_maternaldiet <- read_csv("data/hand1994_maternaldiet.csv")
```

## Question 1 - Slide 15

Make a Two-way table of the hand1994_babies dataset.

```{r}
hand1994_babies # Dataset used
```

```{r}
summary(hand1994_babies) # Dataset summary
```

```{r}
options(digits = 4)
result <- hand1994_babies %>%
  mutate(Weight = cut_interval(Weight, 5)) %>% # Bin Weight in 5 intervals
  count(Survival, Weight) %>%
  contTables(formula = n ~ Weight:Survival, pcRow = T) # Set Weight as exp. variable

result$freqs
```

## Question 4 - Slide 16

Analyze hand1994_elderheight dataset.

```{r}
hand1994_elderheight # Dataset used
```

```{r}
summary(hand1994_elderheight) # Dataset summary
```

```{r}
length(hand1994_elderheight) # Total incidence
```

```{r}
hand1994_elderheight %>% # Binned incidence
  data.frame() %>%
  transmute(Height = cut_width(.,5)) %>% # Bin Height in 5 cm intervals
  count(Height)
```

The Height distribution of the 351 subjects appears to have a normal-like distribution, with mean and median in 160 cm, which is inside the 158 \|-- 162 (here, represented as (158, 162]) bin. The 5 cm interval bin appears to be sufficient to show the distribution.

## Question 6 - Slide 18

Complete the prevalence two-way table of the hand1994_maternaldiet dataset, with control and cases as exp. var. Analyze and discuss.

```{r}
hand1994_maternaldiet # Dataset used
```

```{r}
results <- hand1994_maternaldiet %>%
  pivot_longer(-`Dieta materna`, names_to = "Exp_pop") %>%
  contTables(formula = value ~ Exp_pop:`Dieta materna`, pcRow = T)

results$freqs
```

The group of mothers of babies with spina bifida tend to have a poorer diet in comparison with the control group. However, the control group itself do not have any apparent tendency in diet quality.
