---
title: "2 Means Hypothesis Test List"
author: "Leonardo Yoshiaki Kamigauti"
output:
  pdf_document: default
  html_notebook: default
---

```{r, warning=FALSE, error=FALSE, echo = F, results = 'hide', message=FALSE}
library(tidyverse)
theme_set(theme_minimal())

data <- read_csv("data/treatment.csv") %>%
  mutate(treatment = as.factor(if_else(treatment == 0, "DC", "DBF")))
```

# Assumptions

$$\left\{\begin{matrix}
\alpha = 5\% \\
H_0: \mu_1 = \mu_2 \\ 
H_A: \mu_1 \neq  \mu_2\end{matrix}\right.
$$

# PDI

## Data description

```{r, echo = F}
data %>% 
  select(pdi, treatment) %>%
  summary()
paste("Total obs.:", nrow(data), sep = " ")
```

```{r, warning=FALSE, echo = F, fig.height=4.5, fig.width=8}
data %>%
  select(pdi, treatment) %>%
  ggplot(aes(x = pdi,
             group = treatment)) +
  geom_histogram(aes(y = ..density..,
                     col = treatment),
                 bins = 10,
                 alpha = 0,
                 position="identity") +
  geom_density(aes(col = treatment,
                   fill = treatment),
               alpha = 0.1,
               linetype = 2) +
  ggtitle("Data distribution",
          subtitle = "Histograms and (continuous) kernel density estimation (dashed)")
```

```{r, warning=FALSE, echo = F, fig.height=4.5, fig.width=8}
data %>%
  select(pdi, treatment) %>%
  ggplot(aes(x = treatment,
             y = pdi,
             fill = treatment,
             col = treatment)) +
  geom_violin(trim = FALSE,
              alpha = 0.1,
              linetype = 2) +
  geom_boxplot(width = 0.07,
               alpha = 0) +
  ggtitle("Data comparison",
          subtitle = "Violin and boxplot")
```

## Is the population distribution normal?

```{r, warning=FALSE, echo = F, fig.height=4.5, fig.width=8}
sw <- data %>%
  select(pdi) %>%
  summarise(W = round(shapiro.test(na.omit(pdi))$statistic, digits = 2),
            p = round(shapiro.test(na.omit(pdi))$p.value, digits = 2))

data %>%
  ggplot(aes(sample = pdi)) +
  stat_qq() +
  stat_qq_line() +
  ggtitle("Data normality",
          subtitle = paste("Shapiro-Wilk's results: W =", sw$W, "p-value =", sw$p, sep = " "))
```

There is no evidence that pdi is not normal (p = 0.13), and it is visually close to a normal distribution.

## Are the variances different?

```{r, echo = F}
data %>%
  select(pdi, treatment) %>%
  car::leveneTest(data = ., pdi ~ treatment, center = "mean")
```

There is no evidence of different variances (p = 0.09).

## Are the means different?

```{r, echo = F}
data %>%
  select(pdi, treatment) %>%
  mutate(row = row_number()) %>%
  pivot_wider(names_from = treatment, values_from = pdi) %>%
  t.test(data = ., x = na.omit(.$DC), y = na.omit(.$DBF), alternative = "two.sided", var.equal = T)
```

There is evidence that the distr. means are different (p = 0.03).

# MDI

## Data description

```{r, echo = F}
data %>% 
  select(mdi, treatment) %>%
  summary()
paste("Total obs.:", nrow(data), sep = " ")
```

```{r, warning=FALSE, echo = F, fig.height=4.5, fig.width=8}
data %>%
  select(mdi, treatment) %>%
  ggplot(aes(x = mdi,
             group = treatment)) +
  geom_histogram(aes(y = ..density..,
                     col = treatment),
                 bins = 10,
                 alpha = 0,
                 position="identity") +
  geom_density(aes(col = treatment,
                   fill = treatment),
               alpha = 0.1,
               linetype = 2) +
  ggtitle("Data distribution",
          subtitle = "Histograms and (continuous) kernel density estimation (dashed)")
```

```{r, warning=FALSE, echo = F, fig.height=4.5, fig.width=8}
data %>%
  select(mdi, treatment) %>%
  ggplot(aes(x = treatment,
             y = mdi,
             fill = treatment,
             col = treatment)) +
  geom_violin(trim = FALSE,
              alpha = 0.1,
              linetype = 2) +
  geom_boxplot(width = 0.07,
               alpha = 0) +
  ggtitle("Data comparison",
          subtitle = "Violin and boxplot")
```

## Is the population distribution normal?

```{r, warning=FALSE, echo = F, fig.height=4.5, fig.width=8}
sw <- data %>%
  select(mdi) %>%
  summarise(W = round(shapiro.test(na.omit(mdi))$statistic, digits = 2),
            p = round(shapiro.test(na.omit(mdi))$p.value, digits = 2))

data %>%
  ggplot(aes(sample = mdi)) +
  stat_qq() +
  stat_qq_line() +
  ggtitle("Data normality",
          subtitle = paste("Shapiro-Wilk's results: W =", sw$W, "p-value > 0.005", sep = " "))
```

There is evidence that mdi is not normal (p > 0.005), and it is visually skewed.

## Are the variances different?

```{r, echo = F}
data %>%
  select(mdi, treatment) %>%
  car::leveneTest(data = ., mdi ~ treatment, center = "mean")
```

There is no evidence of no different variances (p = 0.53).

## Are the means different?

```{r, echo = F}
data %>%
  select(mdi, treatment) %>%
  mutate(row = row_number()) %>%
  pivot_wider(names_from = treatment, values_from = mdi) %>%
  t.test(data = ., x = na.omit(.$DC), y = na.omit(.$DBF), alternative = "two.sided", var.equal = F)
```

This t-test considers two different variances. Ignoring the lack of data normality, there is no evidence that the distr. means are different (p = 0.21).

# Conclusion

The sample mean PDI of DBF patients is higher than the sample mean PDI of DC patients.

However, we cannot state that sample mean MDI of DBF patients is different than the sample mean MDI of DC patients.